---
title: "coding_exercice"
author: "Gomina"
date: "2025-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("set up.R")
```

load data
```{r}
data("knz_bison")
```

```{r}
Bisondata <- knz_bison
```

1- Create a new column in the bison dataset that calculates the individual’s age at the time of observation.

```{r}
Bisondata <-Bisondata %>%
  mutate(age= rec_year-animal_yob)
```

2- Write a function that will first filter the bison dataset based on any user-specified weight threshold (returning the filtered dataset).



```{r}
Bisonweight <- function( knz_bison, new_weight) {
   knz_bison %>%
    fiilter (animal_weight >= new_weight)
}
```


3- Create a boxplot showing weight distribution between male and female bison, and then conduct a t-test to see if weight is significantly difference between males and females.




```{r}
ggplot(Bisondata, aes(x = animal_sex, y = animal_weight, fill = animal_sex)) +
  geom_boxplot(alpha = 0.7, outlier.color = "red") +
  labs(
    title = "Weight Distribution Between Male and Female Bison",
    x = "Sex",
    y = "Weight (kg)"
  ) +
  theme_minimal()

# Two sample t-test 
t_test_result <- t.test(animal_weight ~ animal_sex, data = Bisondata)

print(t_test_result)
```



